 

### 2021-04-08 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Vertical profiles in microclimate (Fig. `r fig_NEON_vertical`), leaf traits (see following section), and *evapotranspiration* (*where is this discussed? and do we say anything about evaporation?*) together shape in-situ leaf temperatures according to the principles outlined above (Fig. `r fig_leaf_T`).** 
While the high-light environment of the upper canopy would tend to dramatically increase $T_{leaf}$ above $T_{air}$, particularly under low-wind conditions, strong vertical gradients in leaf morphology and sometimes *evapotranspiration* have a counteractive effect, such that vertical trends in $T_{leaf}$ are much less pronounced than would be expected based on the physical setting alone.
In open-canopy pine forests, light permeating through canopy layers can influence understory leaf temperatures, such that even with understories maintaing $T_{air}$ lower than above canopy $T_{air}$, lower canopy $T_{leaf}$ can be substantially higher than above canopy, even in moist conditions [@mullerEvidenceEfficientNonevaporative2021, @martin_boundary_1999]  
Under some conditions $T_{leaf}$ tracks  $T_{air}$ and therefore differs little between understory and canopy (Fig. `r fig_NEON_vertical`).
For instance, during the dry season in a semi-decidudous tropical moist forest in Panama [@rey-sanchez_spatial_2016]. Sun leaves can also be  cooled relative to $T_{air}$ more than shade leaves [@rey-sanchez_spatial_2016]. 
At night when solar radiation is not a factor, differences in $T_{leaf}$ across the vertical gradient are very modest; for instance, @bolstad_foliar_1999 observed similar nighttime $T_{leaf}$'s in the canopy and understory	of a temperate broadleaf forest.

@@

Interaction between degrees of shade x ontogeny x height with rising temperature needs more scientific attention because a huge part of understory vegetation lies within variations of light and shade spatio-temporal gradient. Leaf traits vary more on the basis on light environment than ontogenetically [REF]

>>>>

**Vertical profiles in microclimate (Fig. `r fig_NEON_vertical`), leaf traits (see following section), and *evapotranspiration* (*where is this discussed? and do we say anything about evaporation?*) together shape in-situ leaf temperatures according to the principles outlined above (Fig. `r fig_leaf_T`).** 
While the high-light environment of the upper canopy would tend to dramatically increase $T_{leaf}$ above $T_{air}$, particularly under low-wind conditions, strong vertical gradients in leaf morphology and sometimes *evapotranspiration* have a counteractive effect, such that vertical trends in $T_{leaf}$ are much less pronounced than would be expected based on the physical setting alone.
In open-canopy needle-leaf forests, light permeating through canopy layers can influence understory leaf temperatures, such that even with understories maintaing $T_{air}$ lower than above canopy $T_{air}$, lower canopy $T_{leaf}$ can be substantially higher than above canopy, even in moist conditions [@mullerEvidenceEfficientNonevaporative2021, @martin_boundary_1999, @zweifelMiddayStomatalClosure2002]  
Under some conditions $T_{leaf}$ tracks  $T_{air}$ and therefore differs little between understory and canopy (Fig. `r fig_NEON_vertical`).
For instance, during the dry season in a semi-decidudous tropical moist forest in Panama [@rey-sanchez_spatial_2016]. Sun leaves can also be cooled relative to $T_{air}$ more than shade leaves [@rey-sanchez_spatial_2016]. 
At night when solar radiation is not a factor, differences in $T_{leaf}$ across the vertical gradient are very modest; for instance, @bolstad_foliar_1999 observed similar nighttime $T_{leaf}$'s in the canopy and understory	of a temperate broadleaf forest.

@@

Interaction between degrees of shade x ontogeny x height with rising temperature needs more scientific attention because a huge part of understory vegetation lies within variations of light and shade spatio-temporal gradient. Leaf traits vary more on the basis on light environment than ontogenetically [@liuPlantSizeEffects2010]
